\chapter{Evaluation}

\section{Requirement Fulfilment}

Overall, all of the requirements that were identified at the beginning of the project have been met to a good standard. The website is able to visualise simple databases and if there are any flaws detected it will advise the user on how to fix these issues. 

Despite the SQL parsing not being as rigorous as expected, the validation in the form works and will catch simple syntax errors that the user might make and inform them of it. The website has a simple and intuitive user interface, with minimalistic styling and theme.

\section{Identification of Requirements}

The identification of requirements was reasonable for the length of this project, the goals were set to be challenging yet attainable. As mentioned in the scope of this project \textit{\hyperref[subsec:scope]{Scope}}, parsing DQL statements and creating an algorithm to automatically repair databases would be outside of the scope of this project. All of the current requirements are necessary for this application to be useful to SQL learners.

\section{Evaluation of Methodology Used}

As discussed in \textit{\hyperref[sec:process]{Chapter 1}}, a process adapted from Scrumban was used. The use of this methodology was beneficial because it gave a solid structure to the development of this project. The weekly sprints were useful because planning was done at the beginning of the sprint for what was going to be worked on in that week. 

This made it easier to focus on a few issues or implementations at a time and the Kanban board allowed the visualisation of issues and their completion stage. During each sprint additional issues were raised as work was being done, these were added to the sprint backlog which was reviewed at the end of the week to create a new set of issues to work on during the future weeks. 

However an improvement for this approach to the project would have been focusing more on testing in the beginning stages of development. The testing framework should have been looked into during spike work and tests should have been written throughout the development process. A framework to write integration tests for the entire website should have also been looked into, it would be useful to be able to automatically test the website as a whole and test the interaction of user interface elements.

\section{Were the design decisions correct?}

The design of the project was mostly correct, more thought might have gone into designing a more efficient parsing framework however, this would require more time than the project allows. A more in-depth parsing framework would allow to have more exhaustive parsing and better syntax error feedback to the user with the line number where the error is and maybe a link to the highlighted text.

The decision to use object-orientated JavaScript was a good design decision since it worked well in this application and the code is readable and structured. Inheritance was used to prevent code duplication in the problem classes. The choice to structure classes to follow the data model of a database was correct since it gave a logical layout to the classes, and gave more context to some of the functions.

The design decisions for the user interface were good because the website is easy to use for SQL learners and conveys the information efficiently and directly. The decision to layout the tables into a tree structure was correct because it was much clearer than laying out the tables in creation order and was appropriate for this application.

\section{Choice of Tools and Development Environment?}

The choice of the development environment was good as there were no issues during the development of the project. The website was developed on a local development server which provided instant feedback on the changes made to either the code or the styling of the website, this was done using Node.js. There were also no issues with missing functionality from the IDE or the website hosting servers. 

The choice of using native JavaScript was good for the current state of the project. In the future if the parsing framework was expanded and more functionality was added to the website which would require more user interface elements then potentially a JavaScript framework would be more appropriate. 

One of these frameworks could be React which would allow the creation of reusable components that would need to be created if the website had more features that would require user interface components. This would require more time to implement these reusable components and learning the framework however it would save time in the future by being able to re-use components. 

The choice of using Bootstrap as the front-end toolkit was correct since it allowed the focus of the project to be on the parsing and the visualising of the tables and less on the styling and design of the website. A more appropriate toolkit might be available that does not focus so much on the responsive and mobile-first design however, for the current state of the project it is satisfactory. 

\subsection{Familiarity of the Technology Used}

The familiarity with web development was adequate with native HTML, CSS and JavaScript which was useful for the project since time did not need to be used to learn these technologies. 

However, Bootstrap was not used before this project and some time had to be used to get used to designing a website using Bootstrap. If some other toolkit was used that was not so focused on responsive design then the re-designing of the \textit{\hyperref[fig:squash]{tables}} might have not been needed. This might have also been avoided if there was familiarity with Bootstrap before the start of this project. 

\section{Lessons Learnt}

If this project would be reworked then the issue of adequate SQL parsing should be addressed. This project should either adapt an already implemented parser from an API and use it to just parse DDL statements. 

Another way to do this is to contribute to a parsing framework that is not fully developed, and add functionality to parse DDL statements. After this issue is solved then more interesting features can be worked on and added to this project.

Deeper exploration of different combinations of frameworks could be done to see if it would be useful when building a more complex website application with more content and features. 

\section{Future Work}

If this project were to be worked on in the future, more advanced detection of flaws in the parsed database should be looked into. This could involve possibly detecting if the database has been normalised and if not see if there is a way to prompt the user to do so. 

Another flaw detection is to see if the table or column name is plural since it is good design to have names be singular, in addition to this it could also check if the correct naming convention is used in names. A feature that could also be implemented is to prompt the user to inform them that a deprecated feature is used, even though it is correct syntax it should be avoided to use deprecated flags.

In addition to more advanced detection of these flaws, there should be some ranking system created to quantify how serious of a flaw it is to the integrity of the database. This could involve ranking a missing primary key as a serious flaw that is high risk where as using a deprecated flag is not even a flaw that has to be fixed, it could be a warning to the user.

Finally, since this project is also aimed to be used in teaching, in the future there should be functionality for people marking student databases. This would involve being able to upload many SQL files at a time, to parse and detect flaws for all of them and output the results in a suitable form. This most likely not be a replacement for marking since the parsing and flaw detection will need to be much more sophisticated for this use case.